---
title: "Assignment2"
author: "zhaoxiuping"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Get a subset with STATE_CODE 6 and SHRP_ID starting with 050:

```{r }
library("dplyr")
library("ggplot2")
library("tidyr")
data2 <- read.csv("/Users/zhaoxiuping/R_zxp/data-assignment/LTPP/iri.csv")
attach(data2)
# data2部分数据展示
sampled_data <- slice_sample(data2, n=10)
sampled_data
```


```{r }
subset_data <- data2 |>
  subset(stringr::str_detect(SHRP_ID,"050") & STATE_CODE==6)
  # |>
  # slice_sample(n=4)
# subset_data部分数据展示
sampled_data <- slice_sample(subset_data, n=10)
sampled_data
```

Obtain the summary statistics of IRI of each section: min, max, and mean:

```{r }
subset_data2 <- data2 |>
  group_by(STATE_CODE,SHRP_ID) |>
  summarise(
    observation = n(),
    # 填补缺失值
    iri_max = max(IRI,na.rm = TRUE),
    iri_min = min(IRI,na.rm = TRUE),
    iri_median = median(IRI,na.rm = TRUE),
    iri_mean = mean(IRI,na.rm = TRUE)
  ) |>
ungroup()
#部分数据展示
sampled_data2 <- slice_sample(subset_data2, n=10)
sampled_data2
```
Sort the summarized data by the averaged IRI in a descending order (report results for one section only):

```{r }
subset_data3 <- arrange(subset_data2,desc(iri_mean))
# 部分数据展示
sampled_data3 <- slice_sample(subset_data3, n=1)
sampled_data3
```

Generate a scatter plot for the averaged IRI against the time for a selected section, and then give your interpretation of the plot: • HINT1:meanIRIvs.date
• HINT2:STATE_CODEandSHRP_IDtogethertoformaprimarykeythatuniquelyidentifiesasection
``` {r }
# STATE_CODE=6 SHRP_ID 以050 开始的section 作图，包含图表
subset_data4 <- subset_data |>
  group_by(VISIT_DATE,STATE_CODE,SHRP_ID) |>
  summarise(
    iri_mean = mean(IRI,na.rm = TRUE)
  ) |>
ungroup()
subset_data5 <- subset_data4 |>
  group_by(VISIT_DATE,STATE_CODE,SHRP_ID) |>
  separate(VISIT_DATE, c("VISIT_DATE", NULL), sep=",") |> mutate(
    VISIT_DATE = as.Date(VISIT_DATE, "%m/%d/%y"), 
    iri_mean = iri_mean
  )

data <- data.frame(
  VISIT_DATE = subset_data5$VISIT_DATE,
  IRI_MEAN = subset_data5$iri_mean,
  STATE_CODE = subset_data5$STATE_CODE,
  SHRP_ID = subset_data5$SHRP_ID
)

scatter_plot <- ggplot(data, aes(x = VISIT_DATE, y = IRI_MEAN)) +
  geom_point(aes(color = paste(STATE_CODE, SHRP_ID)), size = 2) +
  labs(x = "Date", y = "Average IRI", title = "Scatter Plot of meanIRIvs.date") +
  labs(color = "section",size = 2)+
  theme() 
print(scatter_plot)

```

```{r }
#全部的section作图，无图标，图标太多无法显示完全图
subset_data6 <- data2 |>
  group_by(VISIT_DATE,STATE_CODE,SHRP_ID) |>
  summarise(
    iri_mean = mean(IRI,na.rm = TRUE)
  ) |>
ungroup()
subset_data7 <- subset_data6 |>
  group_by(VISIT_DATE,STATE_CODE,SHRP_ID) |>
  separate(VISIT_DATE, c("VISIT_DATE", NULL), sep=",") |> mutate(
    VISIT_DATE = as.Date(VISIT_DATE, "%m/%d/%y"), 
    iri_mean = iri_mean
  )

data <- data.frame(
  VISIT_DATE = subset_data7$VISIT_DATE,
  IRI_MEAN = subset_data7$iri_mean,
  STATE_CODE = subset_data7$STATE_CODE,
  SHRP_ID = subset_data7$SHRP_ID
)

scatter_plot <- ggplot(data, aes(x = VISIT_DATE, y = IRI_MEAN)) +
  geom_point() +
  labs(x = "Date", y = "Average IRI", title = "Scatter Plot of meanIRIvs.date") +
  labs(color = "section",size = 2)+
  theme() 
print(scatter_plot)
```


Get the intersection of the datasets accident and person:

```{r }
accident <- data.frame(read.csv("/Users/zhaoxiuping/R_zxp/data-assignment/CRSS/ACCIDENT.csv"))
person <- data.frame(read.csv("/Users/zhaoxiuping/R_zxp/data-assignment/CRSS/PERSON.csv"))
inter <- intersect(colnames(accident),colnames(person))

inter_ap <- inner_join(
  x=accident,
  y=person,
  by = inter
)
# inter_ap部分数据展示
sampled_inter_ap <- slice_sample(inter_ap, n=5)
sampled_inter_ap
```

Tabulate the total number of observations in each injury severity (INJ_SEV)
• HINT:usesummarise()andgroup_by()

```{r }
observations_injury_severity  <- inter_ap |>
  group_by(INJ_SEV) |>
  summarise(
    count = n()
  )
observations_injury_severity 
```
Merge the accident dataset with the vehicle dataset, and report the dimension of your results and number of missing values in one variable of the right dataset
• HINT:left_join()


```{r }
vehicle <- read.csv("/Users/zhaoxiuping/R_zxp/data-assignment/CRSS/VEHICLE.CSV")
inter_2 <- intersect(colnames(accident),colnames(vehicle))
left_av <- left_join(
  x=accident,
  y=vehicle,
  # by = c("CASENUM","PSU")
  by = inter_2
) |>
  distinct()
#右侧数据集的缺失值
missing_VEH_NO <- sum(is.na(left_av$VEH_NO))
print(missing_VEH_NO)
#合并后的数据集维度
dim(left_av)
```
